---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Rowan Kemmerly"
date: "3.4.23"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
# install.packages("devtools")
devtools::install_github("rstudio-education/dsbox")


library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")

dn <- dennys
lq <- laquinta
```

### Exercise 1

```{r dn-alaska}

dn_ak <- dn %>%
  filter(state == "AK")

nrow(dn_ak)

```

There are three Denny's locations in Alaska. 

### Exercise 2

```{r lq-alaska}

lq_ak <- lq %>%
  filter(state == "AK")
nrow(lq_ak)

```

There are two La Quinta locations in Alaska.

### Exercise 3

There are six possible pairings of Denny's locations and La Quinta locations in Alaska (3 Denny's x 2 La Quintas = 6 pairings). 

I'm now going to join the Denny's and La Quinta data frames (for Alaska) to create these pairings:

```{r alaska-pairs}

dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak

```

### Exercise 4

There are six observations (6 rows) and 11 variables (11 columns): address.x, city.x, state, zip.x, longitude.x, latitude.x, address.y, city.y, zip.y, longitude.y, and latitude.y.


### Exercise 5

We can use the "mutate" function to add a new variable to a data frame while keeping the existing variables. 

Re-creating the Haversine function for our purposes here:

```{r haversine}

haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}

```

### Exercise 6

Calculating the distances between all pairs of Denny’s and La Quinta locations in Alaska:

```{r alaska-distances}

dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))

dn_lq_ak

```

### Exercises 7 and 8

Calculating the minimum distance between a Denny’s and La Quinta for each Denny’s location in Alaska and looking at summary statistics:

```{r min-dist-alaska}

dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance)) %>%
  arrange(closest) %>%
  mutate(address.x = fct_inorder(address.x))

dn_lq_ak_mindist

summary(dn_lq_ak_mindist)

```

In Alaska, 2/3 Denny's locations are about 5 km away from a La Quinta, and 1 Denny's is only about 2 km from a La Quinta. 

```{r visualizing-alaska-dennys}

ak_dennys <- ggplot(dn_lq_ak_mindist, aes(x= address.x, y = closest)) +
  geom_col(fill = '#deebf7') +
  labs(x = "Denny's locations in Alaska", y = "distance (in km) to closest La Quinta", title = "Distance between Denny's and La Quinta locations in Alaska") + 
  theme_minimal() +
    theme(axis.text.x=element_blank())

ak_dennys

```


### Exercise 9

Repeating the same analysis for North Carolina:

```{r nc-dennys}

dn_nc <- dn %>%
  filter(state == "NC")

lq_nc <- lq %>%
  filter(state == "NC")

dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")

dn_lq_nc <- dn_lq_nc %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))

dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance)) %>%
  arrange(closest) %>%
  mutate(address.x = fct_inorder(address.x))

summary(dn_lq_nc_mindist)

nc_dennys <- ggplot(dn_lq_nc_mindist, aes(x= address.x, y = closest)) +
  geom_col(fill = '#718C00') +
  labs(x = "Denny's locations in North Carolina", y = "distance (in km) to closest La Quinta", title = "Distance between Denny's and La Quinta locations in North Carolina") +
  theme_minimal() +
    theme(axis.text.x=element_blank())

nc_dennys


```



### Exercise 10

Repeating the same analysis for Texas:
```{r tx-dennys}

dn_tx <- dn %>%
  filter(state == "TX")

lq_tx <- lq %>%
  filter(state == "TX")

dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")

dn_lq_tx <- dn_lq_tx %>%
  mutate(distatxe = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))

dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distatxe)) %>%
  arrange(closest) %>%
  mutate(address.x = fct_inorder(address.x))

summary(dn_lq_tx_mindist)

tx_dennys <- ggplot(dn_lq_tx_mindist, aes(x= address.x, y = closest)) +
  geom_col(fill = '#BF5700') +
  labs(x = "Denny's locations in Texas", y = "distance (in km) to closest La Quinta", title = "Distance between Denny's and La Quinta locations in Texas") +
  theme_minimal() +
  theme(axis.text.x=element_blank())

tx_dennys


```




### Exercise 11

Repeating the same analysis for Georgia (where I was born):

```{r ga-dennys}

dn_ga <- dn %>%
  filter(state == "GA")

lq_ga <- lq %>%
  filter(state == "GA")

dn_lq_ga <- full_join(dn_ga, lq_ga, by = "state")

dn_lq_ga <- dn_lq_ga %>%
  mutate(distagae = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))

dn_lq_ga_mindist <- dn_lq_ga %>%
  group_by(address.x) %>%
  summarize(closest = min(distagae)) %>%
  arrange(closest) %>%
  mutate(address.x = fct_inorder(address.x))

summary(dn_lq_ga_mindist)

ga_dennys <- ggplot(dn_lq_ga_mindist, aes(x= address.x, y = closest)) +
  geom_col(fill = '#FEE0D2') +
  labs(x = "Denny's locations in Georgia", y = "distance (in km) to closest La Quinta", title = "Distance between Denny's and La Quinta locations in Georgia") +
  theme_minimal() +
  theme(axis.text.x=element_blank())

ga_dennys


```


### Exercise 12

Of the states I examined, I think the joke holds true most in Texas! (Because it's such a negatively skewed distribution.) Though the distribution in Georgia does look quite similar (there just aren't nearly as many Denny's locations all in all).
